---
title: "Peer Assignment - Reproducible Research"
author: "Smilla Tx"
date: "7 3 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lattice)
```

## Loading and preprocessing the data 

```{r}
data <- read.csv(file.choose())
```

## What is the mean total number of steps taken per day? 

The data includes missing values (NA), therefore these need to be defined

```{r}
data_use <- data[!is.na(data[1]),]
#data_use$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", data_use$interval), format='%H:%M'
```

In order to calculate the total number of steps:
```{r}

stepsByDay <- tapply(data_use$steps, data_use$date, sum, na.rm=TRUE)

```

This histogram shows the number of steps per day: 

```{r echo=FALSE}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

Now we calculate the mean and median of the total number we already calculated: 

```{r}
mean_steps <- mean(stepsByDay$steps)
median_steps <- median(stepsByDay$steps)
```

** The mean of total steps per day is 10766.19 and the median is 10765. **

## What is the average daily activity pattern?

First we calculate the average number of steps per interval:

```{r}
average_steps_per_interval_data_use <- get_steps_per_interval(data_use)
```

This histogram shows the number of steps per interval:

```{r echo=FALSE}
plot(average_steps_per_interval_data_use$interval, average_steps_per_interval_data_use$steps, type = "l", xlab = "interval", ylab = "average steps")
```

To get a better overview of the data, we calculate the interval with the highest step rate:

```{r}
max_steps <-  average_steps_per_interval_data_use$steps ==
            max(average_steps_per_interval_data_use$steps)

max_per_interval <- average_steps_per_interval_data_use[max_steps,]
interval_max <- max_per_interval$interval
```

** The interval with the highest average step rate was 835.**

## Imputing missing values

First we calculate the number of missing values in the dataset:

```{r}
missing_values <- sum(is.na(data$steps))
```

Now the missing values will be replaced with the prior calculated average steps per interval:

```{r}
fill_data <- data
na_indices <- which(is.na(fill_data$steps))
ave_steps <- average_steps_per_interval_data_use 

   for(i in na_indices) {
     interval <- fill_data$interval[i]
     fill_data$steps[i] <- ave_steps[ave_steps$interval ==interval,]$steps
   }
```

Now this new data can be found in the fill_data set. 
Now we created a histogram to look at the data:

```{r echo=FALSE}
ggplot(total_steps_per_day_fill_data, aes(x = day, y = steps)) 
    geom_bar(stat = "identity") 
    ylab("total steps")
```

We repeat the step from above with the new data set to calculate the mean and median:

```{r}
mean_fill<- mean(total_steps_per_day_fill_data$steps)
median_fill <- median(total_steps_per_day_fill_data$steps)
```

** Both the median and the mean have the value 10766.19.**

## Are there differences in the activity patters between weekdays and weekends?

First we add a factor to the data to see when was a weekday and when weekend.

```{r}
weekdays_list <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
fill_data$day_of_week <- weekdays(fill_data$date)
fill_data$type_of_day <-
  factor(fill_data$day_of_week %in% weekdays_list,
         levels = c(FALSE, TRUE),
         labels = c("weekend", "weekday"))
```

We now want to look at the differences with a plot:

```{r}
weekday_data <- fill_data[fill_data$type_of_day == "weekday",]
weekend_data <- fill_data[fill_data$type_of_day == "weekend",]
weekday_steps <- get_steps_per_interval(weekday_data)
weekend_steps <- get_steps_per_interval(weekend_data)
```
```{r echo=FALSE}
weekday_steps$type_of_day <- factor(TRUE, levels = c(FALSE, TRUE), labels = c("weekend", "weekday"))
weekend_steps$type_of_day <- factor(FALSE, levels = c(FALSE, TRUE), labels = c("weekend", "weekday"))
weekday_weekend_steps <- rbind(weekday_steps, weekend_steps)
xyplot(steps ~ interval | type_of_day, data = weekday_weekend_steps, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")
```


